---
title: "Strings & Factors"
output: github_document
date: "2022-10-18"
---

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)
library(p8105.datasets)
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## String vectors


```{r}
string_vec = c("my", "name", "is", "jeff")

str_detect(string_vec, "m")
str_replace(string_vec, "jeff", "Jeff")
```
`str_detect`: Did you find anything in this string that matches what you're looking for?


```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "i think")

str_detect(string_vec, "^i think") # refer to beginning of the line -- use ^
str_detect(string_vec, "i think$") # refer to end of the line -- use $
```
This might be a good time to use a regular expression -- 3 phrases start with 'i think' and one ends with 'i think'


```{r}
string_vec = c(
  "Y'all remember Pres. HW Bush?",
  "I saw a green bush",
  "BBQ and Bushwalking at Molonglo Gorge",
  "BUSH -- LIVE IN CONCERT!!"
  )

str_detect(string_vec, "Bush")
#if you want uppercase B or lowercase b, you need an or statement
str_detect(string_vec, "[Bb]ush")
```


```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

# what is the pattern that looks like a number followed by any letter

str_detect(string_vec, "[0-9]")
str_detect(string_vec, "[0-9][A-Z]")
str_detect(string_vec, "[0-9][a-zA-z]")

```


```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

# we want to search for 711 with any character in between 7 and 11

str_detect(string_vec, "7.11")
# character that indicates 'anything' is a '.'

# what if you wanted to search for an actual '.'

```


```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

# how do you search for a special character that has a deeper meaning in regex

str_detect(string_vec, "\\[")
# the backslash is also a special character so need two of them

str_detect(string_vec, "\\[[0-9]")
```


## Why factors are weird

```{r}
factor_vec = factor(c("male", "male", "female", "female"))
# note: levels show that female is first

as.numeric(factor_vec)

# releveling the factor (putting male first)
factor_vec <- fct_relevel(factor_vec, "male")

```


## NSDUH

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"
table_marj = 
  read_html(nsduh_url) %>% 
  html_table() %>% 
  first() %>% 
  slice(-1)
```

tidy up the NSRH data...

```{r}
marj_df <- table_marj %>% 
  select(-contains("P value")) %>% 
  pivot_longer(
    -State,
    names_to = "age_year",
    values_to = "percent"
  ) %>% 
  separate(age_year, into = c("age", "year"), sep = "\\(") %>% 
  mutate(
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent),
    year = str_replace(year, "\\)", "")
  ) %>% 
  filter(
    !(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West", "District of Columbia"))
  )
```


```{r}
marj_df %>% 
  filter(age == "12-17") %>% 
  mutate(State = fct_reorder(State, percent)) %>% 
  ggplot(aes(x = State, y = percent, color = year)) +
  geom_point() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
  )
```


## Restaurant inspections...

```{r}
data("rest_inspec")
```


```{r}
rest_inspec %>% 
  group_by(boro, grade) %>% 
  summarize(n_obs = n()) %>% 
  pivot_wider(
    names_from = grade,
    values_from = n_obs
  )
```


looking at grades of A/B/C

```{r}
rest_inspec <-
  rest_inspec %>% 
  filter(grade %in% c("A", "B", "C"), boro != "Missing") %>% 
  mutate(boro = str_to_title(boro))

rest_inspec %>% 
  group_by(boro, grade) %>% 
  summarize(n_obs = n()) %>% 
  pivot_wider(
    names_from = grade,
    values_from = n_obs
  )
```

let's find pizza places...

```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>% 
  group_by(boro) %>% 
  summarize(n_pizza = n())


rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>% 
  mutate(boro = fct_infreq(boro)) %>% 
  ggplot(aes(x = boro)) +
  geom_bar()
```


```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>% 
  mutate(
    boro = fct_infreq(boro),
    boro = str_replace(boro, "Manhattan", "The")) %>% 
  ggplot(aes(x = boro)) +
  geom_bar()
```
Notice that we removed the factor reordering that we made using fct_infreq



Could also do:

```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>% 
  mutate(
    boro = fct_infreq(boro),
    boro = fct_recode(boro, "The City" = "Manhattan")) %>% 
  ggplot(aes(x = boro)) +
  geom_bar()
```

